---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bis620.2023

<!-- badges: start -->
[![R-CMD-check](https://github.com/zihanwang00/bis620.2023/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/zihanwang00/bis620.2023/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/zihanwang00/bis620.2023/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/zihanwang00/bis620.2023/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

The goal of bis620.2023 is to create a package that allows for the visualization and a platform that facilitates viewers experience with interacting with large and copious amounts of data. It serves to create an interface where viewers can view the distinct features and attributes of the data, specifically comparing to treatment and condition. We apply some of the following operations: creating a world map of trials, creating a condition keyword search, creating an intervention and condition bar chart, creating a pie chart for outcomes, having an ID count for individual countries, and finally creating a picture of a word cloud. 
Creating a world map of trials allows us to see the distribution of trials across the countries. The intervention and condition keyword search allows viewers to filter and see the brief titles or types of trials that they are interested in. The bar chart of the top 10 conditions allows us to select an intervention type and see a list of conditions that are  associated with studies involving that type of intervention. The pie chart allows us to see the proportion on the types of outcomes. The ID count allows us to see the frequency and distribution of subjects across each of the countries. The word cloud provides a visual diagram of the frequency of words that appear in the study. All of these serve to facilitate the ability of processing and visualizing the data more easily as well as making it more interpretable for the viewer to see as well. 

## Installation

You can install the development version of bis620.2023 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("zihanwang00/bis620.2023")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(bis620.2023)
accel |> 
  head(1000) |> 
  plot_accel()
```

```{r, include = FALSE}
library(dplyr)
filtered_countries <- data.frame(countries) |>
  filter(!removed) |>
  group_by(name) |>
  summarise(n =n()) |>
  rename(ID_count = n)
```

Use function count_country_id to count number of trials in specific country
```{r}
count_country_id("United States", countries)
```



## Test Coverage Report

You can run the following code to see test report of the package.

```{r}
library(covr)
report()
```


```{r}
cov <- package_coverage()
print(cov)
```
