---
title: "Process Explanation"
author: "Zihan Wang"
date: "2023-11-08"
output: html_document
---

## For license under DESCRIPTION
License: library(usethis) -- use_gpl_license(version = 2)
## Loading Package
library(devtools) --> load_all() --> can use functions in the package. 
functions in R folder, desccriptions in man (manual) folder
## Documentation
use_roxygen_md(): another package to generate documentation

Document your data (see 'https://r-pkgs.org/data.html')
#### Generate documentation for the function
```{r}
document() 
```
browse Vignettes for package
```{r}
browseVignettes()
```
## Add package

## NAMESPACE
For package to know which file is exported
## Check if package works as expected
```{r}
check()
```
## Testing
```{r}
use_testthat()
use_test("hello")
test()
```
#### report()
```{r}
report()
```
report() == 100%: for the test, all codes written in the packages has been touched
## Add dataset to package
```{r}
use_data_raw() # generate new directory
usethis::use_data(accel, overwrite = TRUE) # giv data directory, have accel.rda file
```
load("../data/accel.rda", verbose = TRUE)
-- not only load the data, also hold the name
```{r}
accel$X[1] = 0 # change the value will not change the original data table
```

### Other
dplyr:: (double quoted) functions exported
dplyr::: (triple quoted) panel functions
```{r}
dplyr:::add_count__
```


