---
title: "Panitumumab Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Panitumumab Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# lecture 10.23

# install()
library(bis620.2023)
library(lubridate)
library(dplyr)
library(ggplot2)

# data(accel)
# t1 = accel$time[5000]
# accel |> 
#   filter((time >= t1) & (time <= t1 + minutes(5))) |>
#   spec_sig(take_log = TRUE) |>
#   filter(freq <= 3) |>
#   plot_accel(x_var = "freq") +
#     theme_bw()
#   
```

## Introduction


## Analysis1: Actual treatments(ATRT) & other variables as death (DTH) (linear model)
variables: actual treatmemt, patient type, AGE, SEX, B_WEIGHT, B_HEIGHT, RACE (some interactive)
```{r}
ae_sl <- left_join(adae, adsl, by = "SUBJID")
```


```{r}
library(corrplot)
library(survival)
library(survminer)
data(adae)
data(biomark)
data(adsl)

# survival plot
ggsurvplot(survfit(Surv(DTHDY, DTH) ~ ATRT, data = ae_sl), 
           data = ae_sl, conf.int = TRUE)

# for numeric variables 
cor_plot(ae_sl)
# for categorical variables
chi_sq(ae_sl$DTH, ae_sl$PRSURG)
chi_sq(ae_sl$DTH, ae_sl$SEX)
chi_sq(ae_sl$DTH, ae_sl$RACE)

# linear model
dth_lm <- lm_model(ae_sl, DTH ~ -1 + PRSURG + ATRT + AGE + RACE + B_METANM + B_HEIGHT)
summary(dth_lm)

```

## Severity level (AESEVCD) and treatment (ATRT), death (DTH)
Want to look at the correlation between severity level and actual treatment & if distribution of death differs on different severity level
```{r, warning = FALSE}
treat_hist(ae_sl, "AESEVCD", "ATRT")
treat_hist(ae_sl, "DTH", "AESEVCD")

# test if two variables are correlated
chi_sq(ae_sl$AESEVCD, ae_sl$ATRT)
chi_sq(ae_sl$AESEVCD, ae_sl$DTH)

# for specific severity
ae_sl %>% 
  filter(AESEVCD == 4) %>% 
  treat_hist("DTH", "ATRT")

ae_sl %>% 
  filter(AESEVCD == 3) %>% 
  treat_hist("DTH", "ATRT")

# check interactive effects
dth_lm2 <- lm_model(ae_sl, DTH ~ -1 + ATRT * AESEVCD)
summary(dth_lm2)

```

## test if biomark is related to DTH, if adding patient types to DTH makes a better prediction

It was hypothesized that patients in the study with KRAS mutations in exons 2, 3, and 4 would have poorer outcomes.

A patient will be considered “Mutant” if there is at least one “Mutant” biomarker in KRAS exons 2, 3, 4. Patients will be considered “Wild-type” if they are not “Mutant” and they have more “Wild-type” markers than “Unknown” or “Failure”. (HW2-Q3)

```{r}
bio <- biomark %>% 
  clean_biomark()
```


```{r}
# exploring biomark types
long_data <- bio_long(bio)
bar_plot(long_data, "Type", "Total")
```


```{r}
# exploring patient types
bar_plot(bio %>% count(patient_type) , "patient_type", "n")
```


```{r}
all <- left_join(ae_sl, bio, by = "SUBJID")
                 
# test if patient type is prognostic with respect to the death outcome
chi_sq(all$DTH, all$patient_type)
```


```{r}
# test if variables are predictive with respect to the death outcome
dth_lm3 <- lm_model(all, DTH ~ -1 + ATRT + patient_type +  patient_type:ATRT)
summary(dth_lm3)
```


```{r}
dth_lm4 <- lm_model(all, 
                    DTH ~ -1 + PRSURG + ATRT + AGE + RACE + B_METANM + 
                      B_HEIGHT + patient_type +  patient_type:ATRT)
summary(dth_lm4)
```

## Conclusion 

